Binary companions of red giant stars with APOGEE: \\ catalog and companions that survive the common envelope

### 2018-02-07 paper/1-catalog/twoface-catalog.tex
More methods

% define macros for math
\newcommand{\hyperpars}{\gamma}
\newcommand{\pars}{\theta}

@@

% Notebook: "Numbers after cuts"
\begin{figure}[hp]
\begin{center}
\includegraphics[width=\textwidth]{nvisits.pdf}
\end{center}
\caption{%
\todo{Update with new apogee db file}
Number of \apogee\ stars in logarithmic bins of number of visits that pass the
quality cuts described in \sectionname~\ref{sec:data}.
In total, this work uses \nvisits\ visits and \nstars\ unique sources.
\label{fig:nvisits}
}
\end{figure}

@@

We want to compute the likelihood of all data from all $N$ stars, $\{D_n\}$,
given a new set of hyper-parameters $\bs{\alpha}$
\begin{equation}
    p(\{D_n\} \given \bs{\alpha}) = \prod_n^N p(D_n \given \bs{\alpha})
\end{equation}
where in the above, we have assumed that this likelihood is separable ( the data
for each source are independent).
The per-source marginal likelihood in the above expression is given by
\begin{align}
    p(D_n \given \bs{\alpha}) &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \,
      p(\bs{w}_n \given \bs{\alpha})\\
    &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \, p(\bs{w}_n \given \bs{\alpha}) \,
      \frac{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}\\
    &= p(D_n \given \bs{\alpha}_0) \, \int \dd \bs{w}_n \,
      \frac{p(\bs{w}_n \given \bs{\alpha})}{p(\bs{w}_n \given \bs{\alpha}_0)} \,
      p(\bs{w}_n \given D_n, \bs{\alpha}_0) \label{eq:marglike} \quad .
\end{align}
Using the Monte Carlo integration approximation, \eqname~\ref{eq:marglike} can
be simplified to a sum over prior value ratios of the $K$ posterior samples for
each $n$ star
\begin{equation}
    \approx \frac{\mathcal{Z}_n}{K_n} \,
      \sum_k^{K_n} \frac{p(y_{nk} \given \bs{\alpha})}{p(y_{nk} \given \bs{\alpha}_0)}
      \quad .
\end{equation}

Factor y from theta, so most priors cancel. Up to multiplicative constant all we need. Above just for one source, full population:

\begin{eqnarray}
    p(\{D_n\} \given \alpha) &\propto& \prod_n^N \frac{1}{K_n} \,
      \sum_k^{K_n} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)}\\
    \ln p(\{D_n\} \given \alpha) &\propto& \sum_n^N \left[
      \ln\left( \sum_k^{K_n} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)} \right)
      - \ln K_n\right]\\
    &\propto& \sum_n^N \left[
      \underset{k}{\rm logsumexp}\left[ \ln{p(y_{nk} \given \alpha)} - \ln{p(y_{nk} \given \alpha_0)} \right]
      - \ln K_n\right]
\end{eqnarray}

TODO: apply to all stars with $>10$ visits

>>>>

@@

% Notebook: "Numbers after cuts"
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.7\textwidth]{nvisits.pdf}
\end{center}
\caption{%
\todo{Update with new apogee db file}
Number of \apogee\ stars in logarithmic bins of number of visits that pass the
quality cuts described in \sectionname~\ref{sec:data}.
In total, this work uses \nvisits\ visits and \nstars\ unique sources.
\label{fig:nvisits}
}
\end{figure}

@@

We want to compute the likelihood of all data from all $N$ stars, $\{D_n\}$,
given a new set of hyperparameters $\bs{\alpha}$
\begin{equation}
    p(\{D_n\} \given \bs{\alpha}) = \prod_n^N p(D_n \given \bs{\alpha})
\end{equation}
where in the above, we have assumed that this likelihood is separable ( the data
for each source are independent).
The per-source marginal likelihood in the above expression is given by
\begin{align}
    p(D_n \given \bs{\alpha}) &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \,
      p(\bs{w}_n \given \bs{\alpha})\\
    &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \, p(\bs{w}_n \given \bs{\alpha}) \,
      \frac{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}\\
    &= p(D_n \given \bs{\alpha}_0) \, \int \dd \bs{w}_n \,
      \frac{p(\bs{w}_n \given \bs{\alpha})}{p(\bs{w}_n \given \bs{\alpha}_0)} \,
      p(\bs{w}_n \given D_n, \bs{\alpha}_0) \label{eq:marglike} \quad .
\end{align}
Using the Monte Carlo integration approximation, \eqname~\ref{eq:marglike} can
be simplified to a sum over prior value ratios of the $K$ posterior samples in
the log-jitter parameter for each $n$ star
\begin{equation}
    \approx \frac{\mathcal{Z}_n}{K} \,
      \sum_k^{K} \frac{p(y_{nk} \given \bs{\alpha})}{p(y_{nk} \given \bs{\alpha}_0)}
\end{equation}
where we have canceled the other priors (over $\bs{\theta}$), and all
normalization constants appear in the constant scale factor $\mathcal{Z}_n$.

The above expresion gives the marginal likelihood of the velocity data for a single source given new hyperparameters $\bs{\alpha}$.
The full marginal likelihood is then the product of these individual likelihoods
\begin{align}
    p(\{D_n\} \given \alpha) &\propto \prod_n^N \frac{1}{K} \,
      \sum_k^{K} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)}
      \quad .
\end{align}
In practice, we evaluate the log-marginal-likelihood
\begin{align}
    \ln p(\{D_n\} \given \alpha) &\propto \sum_n^N \left[
      \ln\left( \sum_k^{K} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)} \right)
      - \ln K\right]\label{eq:lnlike}\\
    &\propto \sum_n^N \left[
      \underset{k}{\textrm{logsumexp}}\left[ \ln{p(y_{nk} \given \alpha)} - \ln{p(y_{nk} \given \alpha_0)} \right]
      - \ln K\right]
\end{align}
where $\textrm{logsumexp}$ (the log-sum-exp trick) provides a more stable
estimate of the sum in \eqname~\ref{eq:lnlike}.

We use 1,825 stars with $>10$ visits and $\log g < 2$ (to avoid upper RGB stars
that have large intrinsic jitter) and maximize the above likelihood to determine
better hyperparameters for the log-jitter parameter distribution.
\figurename~\ref{fig:infer-jitter} shows the distribution corresponding to the
maximum-likelihood hyperparameters, $\alpha^* = (\mu_y^*, \sigma_y^*) = (9.50,
1.64)$.
\todo{discussion about this}
\todo{some discussion of how this relates to empirical bayes?}

% Notebook: "Infer jitter"
\begin{figure}[h]
\begin{center}
\includegraphics[width=\textwidth]{infer-jitter}
\end{center}
\caption{%
\todo{move notebook to figures path, re-do with "run2" or "run1" samples...}
Inferred prior over log-jitter parameter ($y = \ln s^2$) using posterior samples
for 1,825 lower-RGB stars with $>10$ visits.
Left panel shows the ... Right panel shows the ...
\label{fig:infer-jitter}
}
\end{figure}
